---
title: "Pollination analysis"
author: "Kauthar Ismail"
date: "2025-02-26"
output: html_document
---
 This scripture was made using Rmarkdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This scripture will use the various floral traits including flower colour and UV presence,corolla tube depth,nectar sugar content(NSC) and number of flowers per plot for each cultivar:to determine attractiveness to **bumblebees** and **honeybees**
We looked at 119 cultivars of 8 plant groups namely *Salvia  nemorosa*(SV), *Gaillardia  aristata*(GA), *Delosperma  cooperi*(DL),*Lavandula  angustifolia*(LF), *Lavandula   stoechas*(LV), *Sedum   telephium*(SE), *Perovskia   atriplicifolia*(PV) and *Agastache hybrida*(MP)

#### Load the required libraries 

```{r}
library(tidyverse)
```

#### Read in the raw data or csv's

```{r}
getwd()
setwd("~/GIT files/Floral_Traits_Pollination/Data files")
DL<-read_csv("DL.csv")
GA<-read.csv("GA.csv")
LF<-read.csv("LF.csv")
LV<-read.csv("LV.csv")
MP<-read.csv("MP.csv")
PV<-read.csv("PV.csv")
SE<-read.csv("SE.csv")
SV<-read.csv("SV.csv")
```

#### Clean up data as required by combining the multiple data files into one frame

```{r}
data_frame<-bind_rows(GA,PV,SE,LF,LV,MP,DL,SV)
```

#### Choose which columns to include for possible analysis 

```{r}
data_frame_new<-data_frame %>% select (Variety,Bee,Bee_sd,Bumble,Bumble_sd,NSC,NSC_sd,Corolla,Corolla_sd,Flower,
                                       Flower_sd,Colour)
```

#### Binding the variety name and colour together to produce cultivar description

```{r}
data_frame_new <- data_frame_new %>% mutate(Cultivar = paste(Variety, Colour, sep = " "))
```

#### Convert Cultivar from character to factor 

```{r}
data_frame_new$Cultivar<-as.factor(data_frame_new$Cultivar)
```
  
#### Summary of the data {.tabset}

##### Summary
```{r}
summary(data_frame_new)

```

##### Data frame
```{r}
#| collapse = TRUE
data_frame_new
```
####

#### Convert the data to long format and create a plot showing the differences in nectar sugar content(NSC),corolla tube depth,flower number,honeybee and bumblebee number in the first ten cultivars of *G.Aristata* (GA01-GA10)                     
```{r}
data_frame_new %>% pivot_longer(cols=c("NSC","Corolla","Flower","Bee","Bumble"),
                              names_to="variable",values_to="value") %>% slice_head(n=50) %>%
  ggplot() +
    geom_boxplot(aes(y = value, x = Cultivar)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    facet_wrap(vars(variable), scales = "free")

```


