---
title: "Pollination analysis"
output:
  html_document:
    df_print: paged
---
Done by *Kauthar Ismail*

20 February 2025

This scripture will use the various floral traits including flower colour and UV presence,corolla tube depth,nectar sugar content and number of flower per plot for each cultivar:to determine attractiveness to **bumblebees** and **honeybees**
This script was made using Rmarkdown-R notebook

#I set up the R notebook by knitting to allow viewing in one place 
```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load the required libraries 
```{r echo=TRUE}
library(tidyverse)
```

#Read in the raw data 
```{r}
DL<-read_csv("DL.csv")
GA<-read.csv("GA.csv")
LF<-read.csv("LF.csv")
LV<-read.csv("LV.csv")
MP<-read.csv("MP.csv")
PV<-read.csv("PV.csv")
SE<-read.csv("SE.csv")
SV<-read.csv("SV.csv")
```
#Clean up data as required ( first combine the data in one data frame)
```{r}
data_frame<-bind_rows(GA,PV,SE,LF,LV,MP,DL,SV)
            
```
#Choose which columns to include for possible analysis 
```{r}
data_frame_new<-data_frame %>% select (Variety,Bee,Bee_sd,Bumble,Bumble_sd,NSC,NSC_sd,Corolla,Corolla_sd,Flower,
                                       Flower_sd,Colour)
```

#summary of the new data frame to identify classes
```{r}
summary(data_frame_new)
```

#Binding the variety name of the cultivars and the colour together
```{r}
library(dplyr)
data_frame_new <- data_frame_new %>% mutate(Cultivar = paste(Variety, Colour, sep = " "))

```
#Make character(in this case the cultivar) as factor
```{r}
data_frame_new$Cultivar<-as.factor(data_frame_new$Cultivar)
```

<details>
  <summary>Summary of the data</summary>
 #Looking at the summary of the data to ensure factors and numerics 
```{r}
summary(data_frame_new)
data_frame_new
```
</details>


<details>
  <summary>Creating a visual figure representation </summary>
#Convert the data to long format and create a plot showing the                                      differences in nectar sugar content(NSC),corolla,flower,Bees and bumblebees                                  in the different cultivars 
```{r}
data_frame_new %>% pivot_longer(cols=c("NSC","Corolla","Flower","Bee","Bumble"),
                              names_to="variable",values_to="value") %>% slice_head(n=50) %>%
  ggplot() +
    geom_boxplot(aes(y = value, x = Cultivar)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    facet_wrap(vars(variable), scales = "free")

```




```{r}

```

